---
title: "HS Comparison Report for `r my_sch$Long_Name`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      error = FALSE)
```

<br>

# About This Report

This is a demonstration of an interactive report that can be produced with custom content in  interactive elements.

This report is generated with the open source R programming language, and the specific code for generating this report is accessible in open source at [this link](https://github.com/chapinhall/sch-dash-demo). In addition to offering transparency, having code in open source invites collaborators to comment and contribute, regardless of institutional affiliation.

These data represent statistics from the 2018-19 School Year, sourced at [this link](https://data.cityofchicago.org/Education/Chicago-Public-Schools-School-Profile-Information-/kh4r-387c) from the [City of Chicago Data Portal](https://data.cityofchicago.org/).

<!-- /!\ Also mention the data source, here and in the readme -->

<br>
<br>

# Student Body Characteristics

## Comparison vs CPS High School Averages

```{r pair own calculations with CPS}
my_vs_cps <-
  bind_rows(melt(my_sch, 
                 id.vars = "Long_Name",
                 measure.vars = calc_cols$var,
                 variable.name = "variable"),
            hsch_avg %>% mutate(Long_Name = "CPS High School Average")) %>% 
  merge(calc_cols %>% rename(variable = var),
        by = "variable") %>% 
  mutate(Long_Name = factor(Long_Name, 
                            levels = c(hs, "CPS High School Average")),
         var_label = factor(var_label, 
                            levels = calc_cols$var_label))
```

```{r compare own and cps averages side by side, fig.height = 10}
ggplot(my_vs_cps,
       aes(x = Long_Name,
           y = value,
           fill = Long_Name)) +
  geom_bar(stat = "identity",
           position = "dodge") +
  geom_text(aes(label = percent(value,
                                accuracy = 0.1)),
            vjust = -1,
            size = 5) +
  geom_hline(yintercept = 0,
             color = "black") +
  scale_fill_manual(name = "",
                    values = c("#6495ed", "darkgray")) +
  scale_y_continuous(labels = percent,
                     #limits = c(0, 1),
                     expand = expansion(add = c(0, 1.0))) +
  facet_wrap(~ var_label,
             ncol = 2
             #,scales = "free_y"
             ) +
  labs(title = "Comparison of Own Student Body Characteristics to CPS High School Average") +
  theme_minimal() +
  theme(legend.position = "top",
        strip.text = element_text(size = 12),
        plot.title = element_text(hjust = 0,
                                  size = 14),
        axis.text.x  = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y  = element_blank())
```

### Table of Average Value Comparisons

```{r}
hsch_table <- 
  my_vs_cps %>% 
  dcast(var_label ~ Long_Name,
        value.var = "value") %>% 
  rename(`Student Body Characteristc` = var_label)

datatable(hsch_table,
          rownames = FALSE,
          options = list(dom = "t")) %>% 
  formatPercentage(columns = setdiff(colnames(hsch_table), "Student Body Characteristc"))
```


<br>
<br>

## Distribution Among CPS High Schools  {.tabset .tabset-pills .tabset-fade}

```{r}
# hsch_calc_long <- 
#   hsch_calc %>% 
#   melt(id.vars = "Long_Name",
#        measure.vars = calc_cols$var,
#        variable.name = "var") %>% 
#   merge(calc_cols,
#         by = "var") %>% 
#   .[order(value, 
#           decreasing = TRUE)]
# 
# plot_ly(data = hsch_calc_long,
#         type = 'bar',
#         x = ~Long_Name, #~reorder(Long_Name, value),
#         y = ~value,
#         orientation = "h",
#         text = ~percent(value, accuracy = 1),
#         hoverinfo = 'text',
#         transforms = list(
#           list(
#             type = "filter",
#             target = ~var_label,
#             operation = "=",
#             value = calc_cols$var_label[1]
#           )
#         )
# ) %>% 
#   layout(
#     xaxis = list(title = "",
#                  showticklabels = FALSE),
#     yaxis = list(title = "",
#                  tickformat = ".0%"),
#     updatemenus = list(
#       list(
#         y = 1.2,
#         type = "dropdown",
#         active = 1,
#         buttons = 
#           lapply(calc_cols$var_label, 
#                  function(x) 
#                    list(method = "restyle",
#                         args = list("transforms[0].value", x),
#                         label = x))
#       )
#     )
#   )
```


```{r produce distribution plots, results = "asis"}
for (r in 1:nrow(calc_cols)) {
  my_var       <- calc_cols$var[r]
  my_var_label <- calc_cols$var_label[r]
  cat(glue("\n\n### {my_var_label}\n\n"))
  
  my_plot <- 
    ggplot(hsch_calc[j = my_val := get(my_var)],
           aes(x = reorder(Long_Name, my_val),
               y = my_val,
               fill = ifelse(Long_Name == hs,
                             yes = hs,
                             no = "Other HSs") %>% 
                 factor(levels = c(hs, "Other HSs")))) +
    geom_bar(stat = "identity",
             position = "dodge") +
    geom_hline(yintercept = my_sch[j = get(my_var)],
               color = "#6495ed") +
    scale_y_continuous(labels = percent,
                       breaks = sort(c(labeling::extended(range(hsch_calc$my_val, na.rm = TRUE)[1],
                                                          range(hsch_calc$my_val, na.rm = TRUE)[2], m = 5),
                                       my_sch[j = get(my_var)]))) +
    scale_fill_manual(name = "",
                      breaks = c("Other HSs", hs),
                      values = c("darkgray", "#6495ed")) +
    labs(title = glue("Distribution of % {my_var_label} Across CPS High Schools"),
         subtitle = glue("{hs} is highlighted in blue"),
         x = glue("CPS High Schools in Order"),
         y = "") +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "bottom",
          axis.text.y = element_blank(),
          panel.grid = element_blank())
  # /!\ Add annotation of CPS average
  print(my_plot)
    
  cat("\n\n")
}
```

<br>
<br>

# Map

The map below shows all CPS High Schools as dots. Clicking the dots will produce a pop-up with details. `r hs` is highlighted in blue.

<!-- To dos: 
2. add a crosstalk feature to select down by a given feature. See e.g. https://plotly-r.com/client-side-linking.html
3. build a layer selection, perhaps mutually exclusive, by color of relevant value -- https://stackoverflow.com/questions/50641092/r-leaflet-map-change-legends-based-on-selected-layer-group
-->

```{r prep map data}
hsch_map <- 
  hsch_calc %>% 
  mutate(popup = paste(sep = "<br>",
                       paste0("<b>", Long_Name, "</b>"),
                       paste0("% Low Income: ",        percent(pct_lowinc, accuracy = 1)),
                       paste0("% Special Education: ", percent(pct_sped,   accuracy = 1)),
                       paste0("% English Learners: ",  percent(pct_ell,    accuracy = 1))))
```


```{r produce map}
my_radius <- 8
leaflet(hsch_map) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addCircleMarkers(lng = ~School_Longitude, 
                   lat = ~School_Latitude,
                   label = ~Long_Name,
                   radius = my_radius,
                   stroke = FALSE,
                   color = "black",
                   labelOptions = labelOptions(noHide = FALSE, 
                                               textOnly = FALSE),
                   options = markerOptions(riseOnHover = TRUE),
                   popup = ~popup) %>% 
  addCircleMarkers(lng = my_sch$School_Longitude, 
                   lat = my_sch$School_Latitude,
                   label = ~my_sch$Long_Name,
                   stroke = FALSE,
                   color = "#6495ed",
                   radius = my_radius * 1.5,
                   labelOptions = labelOptions(noHide = FALSE, 
                                               textOnly = FALSE),
                   options = markerOptions(riseOnHover = TRUE),
                   popup = ~popup) %>% 
  setView(lng = my_sch$School_Longitude,
          lat = my_sch$School_Latitude,
          zoom = 12)
```

<br>
<br>

# Questions/Comments?

Write us at [nmader@chapinhall.org](mailto:nmader@chapinhall.org). Find the data and code for this project in [this GitHub repository](https://github.com/chapinhall/sch-dash-demo).
